---
title: "Central American Migration"
author: "Ethan Tenison"
date: "2/6/2020"
output: html_document
---

```{r setup, message=FALSE}
library(ggplot2)
library(dplyr)
library(janitor)
library(ggplot2)
library(viridis)
library(raster)
library(maptools)
library(sf)
library(ggspatial)
library(ggpubr)
library(cowplot)
library(scales)
library(extrafont)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidyr)
windowsFonts(Times=windowsFont("Times New Roman"))
```


# Plot of Central American Migrantion into the United States. 


```{r CA_pop_datacleaning}
options(scipen = 999)
df <- read.csv("census_numbers.csv", stringsAsFactors = FALSE)
df <- clean_names(df)
names(df) <- as.matrix(df[4, ])
df <- df[-1, ]
df <- df[6:238, 1:6]
df <- rename(df, country = "Region and country or area")
df$country <- as.character(df$country)
df <-filter(df,country %in% c("El Salvador", "Honduras", "Guatemala", "Nicaragua"))
df <-rename(
      df,
      "Country" = "country",
      "2000" = "2000*",
      "1990" = "1990*",
      "1980" = "1980*",
      "1970" = "1970*",
      "1960" = "1960*"
   )
df_2010 <- read.csv("./ACS_10_1YR_B05006_with_ann.csv")
names(df_2010) <- as.matrix(df_2010[1,])
df_2010 <- df_2010[-1,]
df_2010 <- clean_names(df_2010)
df_2010 <-dplyr::select(
      df_2010,
      "estimate_americas_latin_america_central_america_nicaragua",
      "estimate_americas_latin_america_central_america_honduras",
      "estimate_americas_latin_america_central_america_guatemala",
      "estimate_americas_latin_america_central_america_el_salvador"
   )
df_2010 <- rename(
      df_2010,
      "Nicaragua" = "estimate_americas_latin_america_central_america_nicaragua",
      "Honduras" = "estimate_americas_latin_america_central_america_honduras",
      "Guatemala" = "estimate_americas_latin_america_central_america_guatemala" ,
      "El Salvador" = "estimate_americas_latin_america_central_america_el_salvador"
   )
df_2010 <-gather(df_2010, Country, population, "Nicaragua":"El Salvador")
df_2010 <- mutate(df_2010, year = "2010")
#df_2010$year <- as.factor(df_2010$year)
df_2010$population <- as.numeric(df_2010$population)
df <- dplyr::select(df, "Country", "1960", "1970", "1980", "1990", "2000")
df <- gather(df, year, population, "1960":"2000")
df$population <- as.numeric(gsub(",", "", df$population))
df <- rbind(df, df_2010)
df$year <- as.numeric(df$year)
```

### Central Americans in the United States

```{r ca_pop_ggplot}
plasma_pal <- c("#7E03A8FF","#0D0887FF", "#F89441FF", "#CC4678FF")
theme_set(theme_bw())
ggplot(data = df, aes(x = year, y = population, group = Country)) +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16)) +
   geom_line(aes(color = Country), size = 1) +
   #geom_point(aes(color = Country)) +
   #ggtitle("Population of foreign born Central Americans in the United States") +
   xlab("Year") +
   ylab("Population") +
   scale_y_continuous(label=comma) +
   scale_color_manual(values = plasma_pal) +
   geom_vline(xintercept = 1998, linetype = "dashed") +
   geom_segment(aes(
      x = 1994,
      y = 1000000,
      xend = 1997.5,
      yend = 1000000
   ),
   arrow = arrow(length = unit(0.5, "cm"))) +
   annotate(
      "text",
      x = 1989,
      y = 970000,
      label = expression("Hurricane Mitch \n", paste("hits Honduras")),
      size = 3.5
   ) +
   
   geom_vline(xintercept = 1979, linetype = "dashed") +
   geom_segment(aes(
      x = 1975,
      y = 500000,
      xend = 1978.5,
      yend = 500000
   ),
   arrow = arrow(length = unit(0.5, "cm"))) +
   annotate(
      "text",
      x = 1970,
      y = 450000,
      label = expression("Salvadoran \n", paste("Civil War begins")),
      size = 3.5
   ) 
ggsave("foreign_born_pop.png", )
```
### HDI Map 


```{r HDI}
library(janitor)
library(ggplot2)
df <- read.csv("Human Development Index (HDI).csv", stringsAsFactors = FALSE)
names(df) <- as.matrix(df[1,])
df <- df[-1,]
df <- clean_names(df)
df$country[df$country == "United States"] <- "United States of America"
coords <- data.frame("long" =c(-90.4,-98.8,-86.5,-86.94,-89.4,-85.3,-86.6,-79.55,-78.56,-77.36),
                "lat" = c(15.3,18.9,17.1,14.78,12.6,12.66,9.0,10.5,23.8,17.12),
                "HDI" = c(0.651,0.767,0.720,0.623,0.667,0.651,0.794,0.795,0.778,0.726))
##Setting up the map
theme_set(theme_bw())
world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
world <- left_join(world, df, by = c("sovereignt"="country"))
world[is.na(world)] <- 0
world$x2018[world$x2018 == ".."] <- "0"
world<-world[!(world$x2018 <.6),]
world$x2018 <- as.numeric(world$x2018)
(ggulf <- ggplot(data = world) +
      geom_sf(aes(fill = x2018)) +
       theme_map() +
      annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
               fontface = "italic", color = "grey22", size = 6) +
      coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
      scale_fill_viridis(option = "plasma", direction = -1) +
      guides(fill = guide_colourbar(title.position = "top",  barwidth = 8)) +
      theme( axis.title.x = element_blank(), 
            axis.title.y = element_blank(),
            panel.background = element_rect(fill = "azure"), 
            plot.caption = element_text(hjust = 1.5), 
            panel.border = element_rect(fill = NA),
            text=element_text(family="Times", face = "bold", size=16),
            legend.position = c(0.03, 0.1),
            legend.direction = "horizontal")+
      geom_text(data = coords, x = coords$long, y = coords$lat, label= coords$HDI, nudge_x = 0.25,
                nudge_y = 0.25, check_overlap = 7, size = 5) +
      annotate(geom = "text", x = -91.7, y = 32.26, label = "0.920", 
               color = "white", size = 5) +
      labs(fill = "HDI"))
      
ggsave("central_america_hdi.png", dpi = "retina")
```


### GDP Perchasing power Parity Map 

```{r gdp_percapita}
library(janitor)
library(ggplot2)
df <- read.csv("Human Development Index (HDI).csv", stringsAsFactors = FALSE)
names(df) <- as.matrix(df[1,])
df <- df[-1,]
df <- clean_names(df)
df$country[df$country == "United States"] <- "United States of America"
df$gdp_percapita[is.na(df$gdp_percapita)] <- 0
df$gdp_percapita <- as.numeric(df$gdp_percapita)
coords <- data.frame("long" =c(-90.4,-98.8,-86.5,-86.94,-89.4,-85.3,-86.6,-79.55,-78.56,-77.36),
                "lat" = c(15.3,18.9,17.1,14.78,12.6,12.66,9.0,10.5,23.8,17.12),
                "HDI" = c("$8,462","$19,845","$8,648","$5,134","$8,332","$5,534","$17,671","$25,554", "$21,017","$9,327"))
##Setting up the map
theme_set(theme_bw())
world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
world <- left_join(world, df, by = c("sovereignt"="country"))
world[is.na(world)] <- 0
world$x2018[world$x2018 == ".."] <- "0"
world<-world[!(world$x2018 <.6),]
world$x2018 <- as.numeric(world$x2018)
(ggulf <- ggplot(data = world) +
      geom_sf(aes(fill = gdp_percapita)) +
      theme_map() +
      annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
                size = 6) +
      coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
      scale_fill_viridis(option = "plasma", direction = -1, labels = dollar_format(prefix="$")) +
      theme( axis.title.x = element_blank(), 
            axis.title.y = element_blank(),
            panel.background = element_rect(fill = "azure"), 
            plot.caption = element_text(hjust = 1.5), 
            panel.border = element_rect(fill = NA), 
            text=element_text(family="Times", face = "bold", size=12),
            )+
      geom_text(data = coords, x = coords$long, y = coords$lat, label= coords$HDI, nudge_x = 1,
                nudge_y = 0.25, check_overlap = T) +
      annotate(geom = "text", x = -91.7, y = 32.26, label = "$62,795", 
               color = "white", size = 4) +
      labs(fill = "GDP Per capita, PPP"))
ggsave("gdp_percapita.png", dpi = "retina")
```

### HDI of Latin American 

*Haiti was removed because it skewed the scale. 

```{r hdi_latin_am}
df <- read.csv("Human Development Index (HDI).csv", stringsAsFactors = FALSE)
names(df) <- as.matrix(df[1,])
df <- df[-1,]
df <- clean_names(df)
df$country[df$country == "United States"] <- "United States of America"
df$country[df$country == "Venezuela (Bolivarian Republic of)"] <- "Venezuela"
df$country[df$country == "Bolivia (Plurinational State of)"] <- "Bolivia"
	
##Setting up the map
theme_set(theme_bw())
world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
world <- left_join(world, df, by = c("sovereignt"="country"))
world[is.na(world)] <- 0
world$x2018[world$x2018 == ".."] <- "0"
world$x2018 <- as.numeric(world$x2018)
world<-world[!(world$x2018 <.6),]
(gamerica <- ggplot(data = world) +
      geom_sf(aes(fill = x2018)) +
      coord_sf(xlim = c(-120.15, -30.12), ylim = c(-60, 45), expand = FALSE) +
      scale_fill_viridis(option = "plasma", direction = -1) +
      theme( axis.title.x = element_blank(), text=element_text(family="Times", face = "bold", size=12),
            axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
            panel.border = element_rect(fill = NA))+
      labs(fill = "HDI"))
ggsave("latin_america_hdi.png")
```
### Price of Commodities

```{r commodities}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)

commodities <- read_excel("CMOHistoricalDataAnnual.xlsx", sheet = "Annual Prices (Real)")
commodities <- dplyr::rename(commodities, "year" = "...1")
commodities <- clean_names(commodities)
commodities <- dplyr::select(commodities, year, coffee_arabica, coffee_robusta, sugar_world, banana_us)
commodities <- dplyr::rename(commodities,  "Coffee Arabica" = "coffee_arabica", "Coffee Robusta" = "coffee_robusta", "Sugar" = "sugar_world", "Bananas" = "banana_us")
commodities <-pivot_longer(
              commodities,
              cols = c(
                      "Coffee Arabica",
                      "Coffee Robusta", 
                      "Sugar",
                      "Bananas"
              ),
              names_to = "Commodity",
        
)
plasma_pal <- c("#7E03A8FF","#0D0887FF", "#F89441FF", "#CC4678FF")
p <- ggplot(commodities, aes(x = year, y=value, color = Commodity)) + 
   geom_line(size = 1, show.legend = FALSE) + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16),
         axis.title.x = element_text(vjust = -2),
           axis.title.y = element_text(vjust = 3.5),
         plot.margin = margin(10, 10, 10, 10)) +
   scale_color_manual(values = plasma_pal) + 
   xlab("Year") +
   ylab("Real Prices (Adjusted for Inflation)") +
   scale_y_continuous(labels = scales::dollar)
p + facet_wrap(~Commodity, scale = "free") + theme(strip.background =element_rect(fill="white", color = "white"), panel.border=element_blank())
ggsave("commodity_prices.png", dpi = 500 )
```


### INM Employees
```{r INM_Employees}
library(lubridate)
ana <- read.csv("INM_Employees_HiredDate - INM_Employees_NoMajor.csv", stringsAsFactors = FALSE)
ana <- ana %>% mutate(count = 1) %>%
        mutate(Date.Hired = as.Date(Date.Hired, "%m/%d/%y")) %>% 
        filter(Date.Hired >= as.Date("2018-12-01")) %>%
        group_by(month=floor_date(Date.Hired, "month")) %>% 
        summarize(count = sum(count)) %>%
        add_row(month = as.Date("2018-12-01"), count = 40)
        
theme_set(theme_bw())
ggplot(data = ana, aes(x = month, y = count)) +
   theme(panel.grid.major = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16),
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
   geom_line(size = 1, color = "#7E03A8FF") +
    scale_x_date(date_breaks = "1 month",
                 date_minor_breaks = "1 month",
                 date_labels = "%b %Y") +
   xlab("Month Hired") +
   ylab("INM Agents") +
   geom_vline(xintercept = as.Date("2019-06-01"), linetype = "dashed") +
   geom_segment(aes(
      x = as.Date("2019-05-10"),
      y = 100,
      xend = as.Date("2019-05-30"),
      yend = 100
   ),
   arrow = arrow(length = unit(0.5, "cm"))) +
   annotate(
      "text",
      x = as.Date("2019-04-10"),
      y = 100,
      label = expression("Garduño is appointed \n", paste("as INM director")),
      size = 3.5
   ) +
   
   
   geom_vline(xintercept = as.Date("2018-12-01"), linetype = "dashed") +
   geom_segment(aes(
      x = as.Date("2018-12-30"),
      y = 55,
      xend = as.Date("2018-12-10"),
      yend = 55
   ),
   arrow = arrow(length = unit(0.5, "cm"))) +
   annotate(
      "text",
      x = as.Date("2019-02-01"),
      y = 50,
      label = expression("López Obrador \n", paste("assumes presidency")),
      size = 3.5
   )
   
   
   
        
ggsave("inm_hiring.png", dpi = 500)
```


### Humanitarian Visa Graph (Suchiate only)

```{r motivo_estancia_autorizada}
library(lubridate)
library(readxl)
windowsFonts(Times=windowsFont("Times"))



Humanitarian_Visa_Graph <- read_excel("Humanitarian Visa Graph.xlsx", 
    sheet = "Sheet2")
Humanitarian_Visa_Graph$date <- ymd(Humanitarian_Visa_Graph$date)
Humanitarian_Visa_Graph <- filter(Humanitarian_Visa_Graph, date >= as.Date("2018-12-01") & date <= as.Date("2019-05-01"))
theme_set(theme_bw())
ggplot(data = Humanitarian_Visa_Graph, aes(x = date, y = count_mea)) +
   theme(
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=20), 
         axis.text.x = element_text(vjust = -1, size = 20),
         axis.title.x = element_text(vjust = -2, size = 20),
         axis.text.y = element_text( hjust = 1, size = 20),
         plot.margin = margin(10, 10, 10, 10),) +
   geom_line(size = 1, color = "#7E03A8FF")+
   scale_y_continuous(label=comma) +
    scale_x_date(date_breaks = "1 month",
                 date_minor_breaks = "1 month",
                 date_labels = "%b %Y") +
   geom_vline(xintercept = as.Date("2018-12-07"), linetype = "dashed", size = 1) +
   geom_vline(xintercept = as.Date("2019-01-28"), linetype = "dashed", size = 1) +
   geom_vline(xintercept = as.Date("2019-04-01"), linetype = "dashed", size = 1) +
   geom_vline(xintercept = as.Date("2019-04-23"), linetype = "dashed", size = 1) +
   xlab("Date") +
   ylab("Humanitarian Visas Issued") +
   #labs(caption = "Source: INM Transparency Request")
ggsave("humanitarian_visas.png", width =14, height = 7,dpi = "retina" )

```


### Humanitarian Visa Graph (Total)

```{r motivo_estancia_autorizada}
library(lubridate)
library(readxl)
windowsFonts(Times=windowsFont("Times"))

total_humanitarian_visas <- read_excel("total_humanitarian visas.xlsx", 
    sheet = "Sheet1")


total_humanitarian_visas$Date <- ymd(total_humanitarian_visas$Date)
#total_humanitarian_visas <- filter(total_humanitarian_visas, Date >= as.Date("2018-12-01") & Date <= as.Date("2019-05-01"))
total_humanitarian_visas <- filter(total_humanitarian_visas, Date >= as.Date("2016-06-01"))
theme_set(theme_bw())
ggplot(data = total_humanitarian_visas, aes(x = Date, y = Number_of_Visas)) +
   theme(
         panel.background = element_blank(),
         axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=24), 
         axis.text.x = element_text(angle = 45, hjust= 1, vjust = 1, size = 24),
         axis.title.x = element_text(size = 24),
         axis.title.y = element_text(vjust = 3, size = 24),
         axis.text.y = element_text( size = 24),
         plot.margin = margin(10, 10, 10, 10)) +
   geom_line(size = 1, color = "#7E03A8FF")+
   scale_y_continuous(label=comma) +
    scale_x_date(date_breaks = "3 month",
                 date_minor_breaks = "3 month",
                 date_labels = "%b %Y") +
   #geom_vline(xintercept = as.Date("2018-12-07"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-01-28"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-04-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-04-23"), linetype = "dashed", size = 1) +
   xlab("Date") +
   ylab("Humanitarian Visas Issued") +
   #labs(caption = "Source: INM Transparency Request")
ggsave("total_humanitarian_visas.png", width =14, height = 7,dpi = 500 )



```
### COMAR Funding 

```{r comar_funding}
library(extrafont)
library(readxl)
windowsFonts(Times=windowsFont("Times"))
Comar_Budget_App <- read_excel("Comar Budget and Applications 2013-2020.xlsx")
Comar_Budget_App <- clean_names(Comar_Budget_App)
Comar_Budget_App <- mutate(Comar_Budget_App, apps = "Applications Submitted", perpes = "Pesos per Applicant")
theme_set(theme_bw())
ggplot(data = Comar_Budget_App, aes(x=year, y=refugee_status_applications_submitted, fill=apps)) + 
   theme(legend.background = element_rect(fill = "transparent"), 
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         legend.title = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16), 
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
         axis.text.y.left = element_text(hjust = 1),
         axis.text.y.right = element_text(hjust = 1),
         axis.title.y.right = element_text( vjust = 2.5),
         axis.title.y.left = element_text(vjust = 2.5),
         axis.title.x = element_text(vjust = -1),
         legend.position = "bottom")+
   geom_bar(stat="identity", width = .5) +
   geom_line(data = Comar_Budget_App, aes(x=year, y=applications_per_peso*8, color = perpes), size = 1,stat="identity") + 
   scale_y_continuous(label = comma, sec.axis = sec_axis(~./8, name = "Pesos per Applicant", label = comma)) +
   scale_x_continuous(breaks = c(2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
   scale_color_manual(NULL, values = "#CC4678FF") +
   scale_fill_manual(values = "#0D0887FF") +
   xlab("Year") +
   ylab("Refugee Applications Submitted") +
    geom_segment(aes(
      x = 2014,
      y = 88000,
      xend = 2013.2,
      yend = 88000
   ),
   arrow = arrow(length = unit(0.3, "cm"))) +
   annotate(
      "text",
      x = 2015.9,
      y = 88000,
      label = "$11,288 pesos per applicant",
      size = 4.5
   )
   
   
ggsave("comar_budget.png", dpi = 500)
```

### Crimes Against Migrants 

```{r crimes_against_migrants}
Complaints_byINM <- read_excel("Complaints_byINM.xlsx")
Complaints_byINM <- clean_names(Complaints_byINM)
theme_set(theme_bw())
ggplot(data = Complaints_byINM) +
   theme(legend.background = element_rect(fill = "transparent"), 
         legend.box.background = element_rect(fill = "transparent"),
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16), 
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
         axis.title.x = element_text(vjust = 0),
         axis.text.y = element_text(hjust = 1.25),
         axis.title.y = element_text(vjust = 2)) +
   geom_line(aes(x=year, y=complaint_number), color = "#0D0887FF", size = 1,stat="identity")+
   geom_point(aes(x=year, y=complaint_number), color = "#0D0887FF")+
   geom_text(aes(x= year, y = complaint_number, label=complaint_number),hjust=.25, vjust=-2.5, size = 4)+
   ylim(500, 725)+
   scale_x_continuous(breaks = c(2014,2015,2016,2017,2018,2019,2020)) +
   xlab("Year") +
   ylab("Number of Complaints") 
   
ggsave("inm_complaints.png", dpi = 500)
```
### Price of Hiring Guide

```{r}
library(readxl)
library(janitor)
library(viridis)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(extrafont)
price_coyote <- read_excel("price of hiring a guide.xlsx")
price_coyote <- clean_names(price_coyote)
price_coyote <- gather(price_coyote, year, value, x2012:x2019)
price_coyote <- rename(price_coyote, measure = x1)
price_coyote$year <- gsub("x", "", price_coyote$year)
price_coyote$year <- as.numeric(price_coyote$year)
price_coyote <- filter(price_coyote, measure == "mean price")
 theme_set(theme_bw())
 ggplot(data = price_coyote) + 
    theme(legend.background = element_rect(fill = "transparent"), 
          legend.key = element_rect(fill = "transparent"),
          legend.title = element_blank(),
          legend.spacing = unit(-1, "lines"),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          text=element_text(family="Times", face = "bold", size=16), 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(hjust = 1.25),
          axis.text.y.right = element_text(hjust = -.25),
          axis.title.x = element_text(vjust = 0),
           axis.title.y = element_text(vjust = 3),
           legend.position = "right")+
     scale_color_manual(values=c("#7E03A8FF", "#F89441FF")) +
     scale_x_continuous(breaks = c(2012,2013,2014,2015,2016,2017,2018,2019)) +
    geom_line(aes(x=year, y=value, color = measure), size = 1,stat="identity")+
    xlab("Year") +
   ylab("Cost of Hiring a Guide") 
  
    
   
ggsave("price_hiring_guide.png", dpi = 500)
```

### Percentage Guide
```{r percentage_guide}
library(readxl)
contratoguia <- read_excel("contratoguia.xlsx", 
    sheet = "Contracting by Year")
contratoguia <- na.omit(contratoguia)
contratoguia <- mutate(contratoguia, perch = as.character(Percentage_Yes))
contratoguia$perch=paste(contratoguia$perch, " %")
theme_set(theme_bw())
ggplot(data = contratoguia) +
   theme(legend.background = element_rect(fill = "transparent"), 
         legend.box.background = element_rect(fill = "transparent"),
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=16), 
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
         axis.text.y = element_text(hjust = 1.25),
         axis.title.x = element_text(vjust = -3),
           axis.title.y = element_text(vjust = 3.5),
         panel.grid.minor = element_blank(),
         plot.margin = margin(10, 10, 10, 10)) +
   
   geom_line(aes(x=Year, y=Percentage_Yes), color = "#CC4678FF", size = 1,stat="identity")+
   geom_point(aes(x=Year, y=Percentage_Yes), color = "#CC4678FF")+
   geom_text(aes(x= Year, y = perc_pos, label=perch),hjust=.35, vjust=-1, size = 4)+
   ylim(40, 65) +
   scale_x_continuous(breaks = c(2011,2012,2013,2014,2015,2016,2017,2018,2019)) +
   xlab("Year") +
   ylab("Hired a Guide (%)") +
ggsave("hired_guide.png", dpi = 500, height = 4.5, width =7)
```




```{r inm_employees}
library(readr)
library(dplyr)
INM_EMPLOYEES_CLEAN <- read_csv("INM_EMPLOYEES_CLEAN.csv")
inm_emp<- INM_EMPLOYEES_CLEAN %>% dplyr::select(State) %>% filter(State != "" ) %>% mutate(number = 1) 
inm_emp<- inm_emp %>% dplyr::group_by(State) %>% summarize(number_agents = sum(number))
adds <- data.frame("State" = c("COLIMA", "MEXICO"), "number_agents" = c(0,0))
inm_emp <- rbind(inm_emp,adds)
                         # Install stringi package
library("stringi")   
#inm_emp$State <- stri_trans_totitle(inm_emp$State) 
library(sf)
mexico <- st_read("mex_admbnda_adm1_govmex.shp")
df_mxstate <- left_join(mexico,inm_emp, by = c("ADM1_ES"="State"))
df_mxstate <- dplyr::select(df_mxstate, ADM1_ES, number_agents, geometry)
```

## Map showing INM Apprehensions

```{r inmmap}

library(readr)
library(dplyr)
library(scales)
INM_EMPLOYEES_CLEAN <- read_csv("INM_EMPLOYEES_CLEAN.csv")
inm_emp<- INM_EMPLOYEES_CLEAN %>% dplyr::select(State) %>% filter(State != "" ) %>% mutate(number = 1) 
inm_emp<- inm_emp %>% dplyr::group_by(State) %>% summarize(number_agents = sum(number))
adds <- data.frame("State" = c("COLIMA", "MEXICO"), "number_agents" = c(0,0))
inm_emp <- rbind(inm_emp,adds)
                         # Install stringi package
library("stringi")   
#inm_emp$State <- stri_trans_totitle(inm_emp$State) 
library(sf)
mexico <- st_read("mex_admbnda_adm1_govmex.shp")
df_mxstate <- left_join(mexico,inm_emp, by = c("ADM1_ES"="State"))
df_mxstate <- dplyr::select(df_mxstate, ADM1_ES, number_agents, geometry)



##Setting up the map
theme_set(theme_bw())

(inm_agents_state <- ggplot(data = df_mxstate) +
      geom_sf(aes(fill = number_agents)) +
      theme_map() +
      scale_fill_viridis(option = "plasma", direction = -1) +
       guides(fill = guide_colourbar(title.position = "top",  barwidth = 10)) +
      theme( axis.title.x = element_blank(), 
            axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
            plot.caption = element_text(hjust = 1.5), 
            panel.border = element_rect(fill = NA),
            text=element_text(family="Times", face = "bold", size=16),
            legend.position = c(0.03, 0.1),
            legend.direction = "horizontal")+
      labs(fill = "Number of Agents"))
ggsave("inm_Agents_state.png", dpi = 500)



```


```{r}

library(readr)
library(tidyr)
library(lubridate)
apps_amlo <- read_csv("INM Apprehensions AMLO Pres.csv")
data_long <- gather(apps_amlo, Date, Apprehensions, "2018-01-01":"2020-02-01", factor_key=TRUE)
data_long$Date <- as.character(data_long$Date)
data_long$Date <- ymd(data_long$Date)
data_long <- data_long %>% filter(Country == "Total")  %>% filter(Date >= as.Date("2018-11-01"))
plasma_pal <- c("#7E03A8FF") # "#0D0887FF", "#F89441FF", "#CC4678FF"
windowsFonts(Times=windowsFont("Times"))

theme_set(theme_bw())
ggplot(data = data_long) +
   theme(
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         legend.position = "none",
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=24), 
         axis.text.x = element_text(hjust= 1, angle = 45),
         axis.title.x = element_text(),
         axis.title.y = element_text(vjust = 2.5)) +
   geom_line(aes(x=Date, y=Apprehensions, color = Country),size = 1,stat="identity")+
   #geom_point(aes(x=Date, y=Apprehensions, color = Country))+
    scale_x_date(date_breaks = "2 month",
                 date_minor_breaks = "2 month",
                 date_labels = "%b %Y") +
    scale_y_continuous(label=comma) +
   scale_color_manual(values = plasma_pal) +  
   #geom_vline(xintercept = as.Date("2019-06-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-10-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2020-01-01"), linetype = "dashed", size = 1) +
   #xlab("Date") +
   #ylab("Apprehensions") 
ggsave("inm_apprehensions_amlo_from2019.png", height = 7, width = 14, dpi = 500)

```


```{r apprehensions_total}

library(readr)
library(tidyr)
library(lubridate)
INM_Apprehensions_total <- read_csv("INM Apprehensions total.csv")
data_long <- gather(INM_Apprehensions_total, Date, Apprehensions, "2001-01-01":"2020-02-01", factor_key=TRUE)
data_long$Date <- as.character(data_long$Date)
data_long$Date <- ymd(data_long$Date)
data_long <- data_long %>% filter(Country == "Total") # %>% filter(Date >= as.Date("2018-11-01"))
plasma_pal <- c("#7E03A8FF") # "#0D0887FF", "#F89441FF", "#CC4678FF"
windowsFonts(Times=windowsFont("Times"))

theme_set(theme_bw())
ggplot(data = data_long) +
   theme(
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         legend.position = "none",
         panel.background = element_blank(), axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=24), 
         axis.text.x = element_text(hjust= 1, angle = 45),
         axis.title.x = element_text(),
         axis.title.y = element_text(vjust = 2.5)) +
   geom_line(aes(x=Date, y=Apprehensions, color = Country),size = 1,stat="identity")+
   #geom_point(aes(x=Date, y=Apprehensions, color = Country))+
    scale_x_date(date_breaks = "1 year",
                 date_minor_breaks = "1 year",
                 date_labels = "%Y") +
    scale_y_continuous(label=comma) +
   scale_color_manual(values = plasma_pal) +  
   #geom_vline(xintercept = as.Date("2019-06-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-10-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2020-01-01"), linetype = "dashed", size = 1) +
   #xlab("Date") +
   #ylab("Apprehensions") 
ggsave("inm_apprehensions_amlo_from2001.png", height = 7, width = 14, dpi = 500)





```


```{r inm_cbp_apprehensions}

library(readr)
library(tidyr)
library(lubridate)

data_long <- read_csv("inm and cbp apprehensions dec 2018 - feb 2020.csv")
data_long$`Migration Enforcement` <- as.factor(data_long$`Migration Enforcement`)

data_long$Month <- as.character(data_long$Month)
data_long$Month <- mdy(data_long$Month)
data_long <- data_long %>% filter(Month >= as.Date("2018-11-01"))
plasma_pal <- c("#F89441FF", "#0D0887FF") # , "#F89441FF", "#CC4678FF"
windowsFonts(Times=windowsFont("Times New Roman"))

theme_set(theme_bw())
ggplot(data = data_long, aes(x=Month)) +
   theme(
         legend.key = element_rect(fill = "transparent"), 
         legend.spacing = unit(-1, "lines"),
         legend.position = "right",
         legend.title = element_blank(),
         panel.background = element_blank(),
         axis.line = element_line(colour = "black"),
         text=element_text(family="Times", face = "bold", size=30), 
         axis.text.x = element_text(hjust= 1, angle = 45),
         axis.title.x = element_text(),
         axis.title.y = element_text(vjust = 2)) +
   geom_line(aes( y=Apprehensions, color = `Migration Enforcement`),size = 1)+
   #geom_point(aes(x=Month, y=Apprehensions, color = Country))+
    scale_x_date(date_breaks = "2 month",
                 date_minor_breaks = "2 month",
                 date_labels = "%b %Y") +
    scale_y_continuous(label=comma) +
   scale_color_manual(values = plasma_pal) +  
   #geom_vline(xintercept = as.Date("2019-06-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2019-10-01"), linetype = "dashed", size = 1) +
   #geom_vline(xintercept = as.Date("2020-01-01"), linetype = "dashed", size = 1) +
   #xlab("Date") +
   #ylab("Apprehensions") 
ggsave("inm_apprehensions_amlo_from2019.png", height = 7, width = 14, dpi = 900)


```




```{r comar_offices}


library(readr)
library(ggmap)

COMAR_offices <- read_csv("COMAR_offices.csv")
COMAR_offices$`Type of Office` <- as.factor(COMAR_offices$`Type of Office`)

theme_set(theme_bw())

  ggplot(data = world) +
    geom_sf() +
    theme_map() +
    geom_point(data = COMAR_offices, shape = 25, aes(y=Latitude, x = Longitude, fill = `Type of Office`), size = 3) +
    scale_fill_viridis_d(option = "plasma", direction = 1) +
    coord_sf(xlim = c(-120.15, -85.12), ylim = c(12, 35), expand = FALSE) +
    geom_text(data = COMAR_offices, x = COMAR_offices$fix_long, y = COMAR_offices$fix_lat, label= COMAR_offices$City,
                nudge_y = -50, check_overlap = 1, size = 3) +
    annotate(
      "text",
      x = -90.70510,
      y = 18.27270,
      label = "Tenosique",
      size = 3
   ) +
    theme( axis.title.x = element_blank(), 
            axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
            plot.caption = element_text(hjust = 1.5), 
            panel.border = element_rect(fill = NA),
            text=element_text(family="Times", face = "bold", size=16),
            legend.position = c(0.03, 0.15),
            legend.title = element_blank()
            )

  ggsave("inm_offices.png", dpi = 500)
```

```{r inm_detention}

library(readxl)
library(ggmap)

inm_facilities_2020 <- read_excel("inm-facilities-2020.xlsx", 
    col_types = c("skip", "text", "text", 
        "text", "text", "numeric", "numeric"))

inm_facilities_2020$facilitytype <- as.factor(inm_facilities_2020$facilitytype)


 
coords <- data.frame("long" =c(-100.133321,-99.014828),# 
                "lat" = c(20.391975,33.285146),#
                "city" = c("Mexico", "United States")) #

theme_set(theme_bw())
plasma_pal <- c("#7E03A8FF","#0D0887FF", "#F89441FF") #"#CC4678FF"

  ggplot(data = world) +
    geom_sf() +
    theme_map() +
    geom_point(data = inm_facilities_2020, aes(y=latitude, x = longitude, shape =facilitytype, color = facilitytype),alpha = 0.8, size = 3) +
    scale_color_manual(values = plasma_pal) +
    coord_sf(xlim = c(-120.15, -85.12), ylim = c(12, 35), expand = FALSE) + ## c(-120.15, -75.12), ylim = c(10, 35),
    #geom_text(data = coords, x = coords$long, y = coords$lat, label= coords$city,
     #           check_overlap = 1, size = 5) +
    theme( axis.title.x = element_blank(), 
            axis.title.y = element_blank(),
           panel.background = element_rect(fill = "azure"), 
            plot.caption = element_text(hjust = 1.5), 
            panel.border = element_rect(fill = NA),
            text=element_text(family="Times", face = "bold", size=16),
            legend.position = c(0.03, 0.15),
            legend.title = element_blank()
            )

  ggsave("inm_detention_centers.png", dpi = 500)

```
